Assessing the Global COVID19 Impact on Air Transport with Open Data

### 2021-07-21 paper.Rmd

The concept of resilience (and robustness) is intimately linked with the construct of disruptions.

@@

Conceptually, resilience R can be measured as the observed loss in quality of service (performance) over the time to recovery, t1-t2, for a certain level of disruption. Thus, mathematically, this represents the area covered by

@@

>>>>

The concept of resilience (and robustness) is intimately linked with the construct of disruptions. Any given disturbance forces a disruption in the level of service. Following Gluchshenko's work (2012), the longer the system takes to recover to a previously unperturbed state, the less resilient the system is. In complement, from Koeller's work (2015), the deeper the level of disruption, also less resilient the system is. As a result, a possible interpretation for resilience metrics is found considering both the duration of the disrupted service performance, and its disruption level - or how much a system level of service is affected.

@@

Consequently, resilience R can be conceptually measured as the observed loss in quality of service (performance) over the time to recovery, t1-t2, for a certain level of disruption. Thus, mathematically, this represents the area covered by

@@

However, an evident question arises from this approach: what is the proper threshold for the service to be considered disrupted? Certain that, for some indicators, this could be a fixed value, this paper proposes an alternative method. The disruption threshold can be viewed as an adptive function of the recent traffic levels. As a result, the disruption threshold level would not be a fixed value, but a variable one, taking into account the recent trends to define a normal band based on moving windows. The example below shows this potential approach.
```{r}
#-------Remember to load and run "x0_data_preparation.Rmd". The object below is produced there!--------#
ResilienceMovingBands +
  theme(legend.position = "right")
```

Instead of a horizontal line (fixed value), the disruption threshold adapts according to the recent trends in traffic. In the figure above, the red line is the disruption threshold, defined as  the `r 100*lowlim`th percentile of the daily traffic, considered the previous `r slide` days. Naturally, those are empirically suggested values that could be adapted according to any particular needs. In addition, the blue line is set to be a potential indicator of disruptions in the opposite direction, i.e. a positive disruption, caused by an abnormal excessive demand.
